---
title: "How Many Local Epochs?"
author: "Keivan Bolouri -- Student ID: 806338280"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
colorlinks: true
linkcolor: blue
urlcolor: blue
citecolor: blue
toccolor: blue
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 4
  word_document:
    toc: true
    toc_depth: '4'
  html_document:
    toc: true
    toc_depth: '4'
    df_print: paged
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage[margin=1in]{geometry}
- \usepackage{xcolor}
- \usepackage{tocloft}
- \renewcommand{\cfttoctitlefont}{\color{blue}\Large\bfseries}
- \renewcommand{\cftsecfont}{\color{blue}}
- \renewcommand{\cftsecpagefont}{\color{blue}}
- \renewcommand{\cftsubsecfont}{\color{blue}}
- \renewcommand{\cftsubsecpagefont}{\color{blue}}
- \renewcommand{\cftdot}{\color{blue}.}
- \usepackage{pdflscape}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = FALSE,
  message   = FALSE,
  warning   = FALSE,
  fig.align = "center",
  fig.width  = 10,
  fig.height = 5
)
library(knitr)
library(kableExtra)
```

---

# Problem Overview

Three data owners each hold a private local dataset $(X_j, y_j)$ with $p = 600$
predictors. They collaborate to estimate **one shared** 600-dimensional Lasso
coefficient vector $\boldsymbol{\beta}$ without sharing raw data.

The roadmap specifies the following node sizes and fixed aggregation weights:

| Node | Training samples | Validation samples | Weight $w_j$ |
|------|:----------------:|:------------------:|:------------:|
| Node 1 | 400 | 100 | 0.50 |
| Node 2 | 240 |  60 | 0.30 |
| Node 3 | 160 |  40 | 0.20 |

The algorithm alternates between:

1. **Local step** — each node $j$ runs $E$ coordinate-descent (CD) epochs of
   Lasso, initialised from the current global consensus vector $\boldsymbol{\beta}_t$.
2. **Aggregation step** — the central server computes
   $\boldsymbol{\beta}_{t+1} = \sum_{j=1}^{3} w_j\,\boldsymbol{\beta}_{j,t}$.

Convergence is declared when
$\|\boldsymbol{\beta}_{t+1} - \boldsymbol{\beta}_t\|_2 < 10^{-5}$.

---

# Data Loading

> **Note on file naming vs. roadmap node order.**  The CSV files are named
> `node1`, `node2`, `node3`, but the actual sample sizes stored in those files
> are (160, 240, 400) for nodes 1, 2, 3 respectively — the *reverse* of the
> roadmap's assignment of (400, 240, 160).  To honour the roadmap specification
> exactly (Node 1 = 400 samples, weight 0.50; Node 2 = 240 samples, weight 0.30;
> Node 3 = 160 samples, weight 0.20), we load the CSVs and then **re-label** so
> that the largest dataset carries the largest weight.

```{r load-data, echo=FALSE, message=FALSE, warning=FALSE}
BASE <- "."

load_csv_node <- function(n, base = BASE) {
  list(
    X_tr  = as.matrix(
               read.csv(file.path(base, paste0("node", n, "_X_train.csv")))),
    y_tr  = as.numeric(unlist(
               read.csv(file.path(base, paste0("node", n, "_y_train.csv"))))),
    X_val = as.matrix(
               read.csv(file.path(base, paste0("node", n, "_X_val.csv")))),
    y_val = as.numeric(unlist(
               read.csv(file.path(base, paste0("node", n, "_y_val.csv")))))
  )
}

raw1 <- load_csv_node(1)   # 160 train / 40  val
raw2 <- load_csv_node(2)   # 240 train / 60  val
raw3 <- load_csv_node(3)   # 400 train / 100 val

# Re-label to match the project specification:
# Node 1 = 400 samples (weight 0.50), Node 2 = 240 (0.30), Node 3 = 160 (0.20)
node1 <- raw3
node2 <- raw2
node3 <- raw1
```

```{r node-summary-table, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)

node_summary <- data.frame(
  Node       = paste0("Node ", 1:3),
  Weight     = c(0.50, 0.30, 0.20),
  Train      = c(nrow(node1$X_tr), nrow(node2$X_tr), nrow(node3$X_tr)),
  Val        = c(nrow(node1$X_val), nrow(node2$X_val), nrow(node3$X_val)),
  Predictors = c(ncol(node1$X_tr), ncol(node2$X_tr), ncol(node3$X_tr))
)

kable(
  node_summary,
  format    = "latex",
  col.names = c(
    "Node",
    "Weight ($w_j$)",
    "Train rows",
    "Val rows",
    "Predictors ($p$)"
  ),
  align    = c("l", "c", "c", "c", "c"),
  caption  = "Dataset dimensions.",
  booktabs = TRUE,
  escape   = FALSE
) |>
  kable_styling(
    latex_options = c("striped", "hold_position"),
    full_width = FALSE
  )
```


---

# Core Functions

## Soft-Threshold Operator

The element-wise soft-threshold (proximal operator for the $\ell_1$ penalty):

$$\text{soft}(\rho,\,\lambda) = \operatorname{sign}(\rho)\,\max(|\rho|-\lambda,\;0)$$

```{r soft-threshold, echo=FALSE}
soft_thresh <- function(rho, lambda) sign(rho) * pmax(abs(rho) - lambda, 0)
```

## One Coordinate-Descent Epoch

For the local Lasso objective
$$\mathcal{L}_j(\boldsymbol{\beta}) =
  \frac{1}{2m_j}\|y_j - X_j\boldsymbol{\beta}\|_2^2
  + \lambda_j\|\boldsymbol{\beta}\|_1,$$
one epoch is one full sequential pass over all $p = 600$ coordinates.

**Update rule for coordinate $k$** (roadmap notation):
$$r_k = y_j - X_j\boldsymbol{\beta} + x_{jk}\beta_k,\quad
  \rho_k = x_{jk}^\top r_k,\quad
  z_k = \|x_{jk}\|_2^2,\quad
  \beta_k \;\leftarrow\; \frac{\operatorname{soft}(\rho_k,\;\lambda_j)}{z_k}.$$

> **Implementation note.** The roadmap writes $\operatorname{soft}(\rho_k,\lambda_j)$
> where $\lambda_j$ is defined on the *normalised* objective scale
> $\frac{1}{2m_j}\|\cdot\|^2$.  Because I define $\lambda_j$ on the normalised
> objective scale, the equivalent coordinate update using the unnormalised residual
> $r_k = y_j - X_j\boldsymbol{\beta} + x_{jk}\beta_k$ uses the threshold
> $\lambda_j m_j$.  Thus the reported $\lambda$ values should be interpreted
> under the normalised-objective convention; the code uses $\lambda_j m_j$
> so that the residual need not be rescaled inside the inner loop.

```{r cd-epoch, echo=FALSE}
cd_epoch <- function(beta, X, y, lam) {
  m <- nrow(X)
  p <- ncol(X)
  z <- colSums(X^2)
  r <- drop(y - X %*% beta)

  for (k in seq_len(p)) {
    r       <- r + X[, k] * beta[k]
    rho_k   <- as.numeric(crossprod(X[, k], r))
    beta[k] <- if (z[k] > 0) soft_thresh(rho_k, lam * m) / z[k] else 0
    r       <- r - X[, k] * beta[k]
  }
  beta
}
```

## Node Update Function

```{r node-update, echo=FALSE}
node_update <- function(node, beta_init, lam, E) {
  b <- beta_init
  for (e in seq_len(E)) {
    b <- cd_epoch(b, node$X_tr, node$y_tr, lam)
  }
  b
}

node_loss_report <- function(node, beta, lam) {
  node_train_loss(beta, node$X_tr, node$y_tr, lam)
}
```

## Loss Functions

```{r loss-functions, echo=FALSE}
node_train_loss <- function(beta, X, y, lam) {
  m     <- nrow(X)
  resid <- drop(y - X %*% beta)
  0.5 / m * sum(resid^2) + lam * sum(abs(beta))
}

# Validation loss = 0.5 * MSE (no penalty) — used for lambda selection only
node_val_loss <- function(beta, X_val, y_val) {
  resid <- drop(y_val - X_val %*% beta)
  0.5 * mean(resid^2)
}
```

## Lambda Tuning (Local, Per-Node)

Each node independently searches a log-scale grid and selects the $\lambda_j$
that minimises its own *validation* loss.  The selected $\lambda_j$ is
**frozen** for all communication rounds.

```{r tune-lambda, echo=FALSE}
LAM_GRID <- exp(seq(log(1), log(1e-4), length.out = 30))

tune_lambda <- function(node, lam_grid, max_iter = 300, tol = 1e-7) {
  best_lam  <- NA_real_
  best_loss <- Inf

  for (lam in lam_grid) {
    beta <- numeric(ncol(node$X_tr))
    for (iter in seq_len(max_iter)) {
      beta_old <- beta
      beta     <- cd_epoch(beta, node$X_tr, node$y_tr, lam)
      if (sqrt(sum((beta - beta_old)^2)) < tol) break
    }
    vl <- node_val_loss(beta, node$X_val, node$y_val)
    if (vl < best_loss) { best_loss <- vl; best_lam <- lam }
  }
  list(lambda = best_lam, val_loss = best_loss)
}
```

---

# Step 1 — Tune $\lambda_j$ at Each Node {#sec-tune}

```{r tune-all, echo=FALSE, cache=TRUE}
set.seed(42)
tune1 <- tune_lambda(node1, LAM_GRID)
tune2 <- tune_lambda(node2, LAM_GRID)
tune3 <- tune_lambda(node3, LAM_GRID)

lam1 <- tune1$lambda
lam2 <- tune2$lambda
lam3 <- tune3$lambda

lam_table <- data.frame(
  Node    = paste0("Node ", 1:3),
  Train_n = c(nrow(node1$X_tr), nrow(node2$X_tr), nrow(node3$X_tr)),
  Val_n   = c(nrow(node1$X_val), nrow(node2$X_val), nrow(node3$X_val)),
  Weight  = c(0.50, 0.30, 0.20),
  Lambda  = round(c(lam1, lam2, lam3), 6),
  Val_MSE = round(c(tune1$val_loss, tune2$val_loss, tune3$val_loss), 4)
)

kable(lam_table,
      col.names = c("Node", "Train $n$", "Val $n$", "Weight $w_j$",
                    "Optimal $\\lambda_j$", "Val. Loss (0.5$\\times$MSE)"),
      align     = c("l","c","c","c","c","c"),
      caption   = "Locally selected regularisation parameters ($\\lambda$ grid search).",
      booktabs  = TRUE) |>
  kable_styling(latex_options = c("striped","hold_position"),
                full_width = FALSE)
```

**Interpretation.**  
Node 1 (largest dataset, 400 samples) and Node 3 (160 samples) select the same
grid value, suggesting that this level of regularisation gives the best
validation performance for both nodes.  Node 2 selects a larger $\lambda$
(stronger shrinkage), indicating that stricter penalisation improves its
held-out fit — possibly due to higher local noise or a sparser signal structure.

---

# Step 2 — Federated Training Loop {#sec-fed}

## Server Loop

The central server stores only $\boldsymbol{\beta}_t$, the fixed weights, and
scalar loss diagnostics.  It calls `node_update()` to obtain local vectors.
For computing the global loss it calls `node_loss_report()`, which encapsulates
the node's data access — mimicking the privacy boundary where each node would
evaluate its own scalar loss at the broadcast $\boldsymbol{\beta}_{t+1}$ and
return **only that scalar** to the server.  In this single-machine simulation
both functions access the node objects directly for convenience, but conceptually
no raw $X_j$ or $y_j$ ever crosses the privacy boundary to the server.

**Global loss definition.**  At round $t$, *after* aggregation, the server
evaluates each node's training loss at the newly formed consensus vector
$\boldsymbol{\beta}_{t+1}$ and records the weighted average:
$$\text{GlobalLoss}(t) = \sum_{j=1}^{3} w_j\,\mathcal{L}_j\!\left(\boldsymbol{\beta}_{t+1}\right).$$
This is the correct global loss at the aggregated model, not at the local
pre-aggregation vectors.

```{r federated-loop, echo=FALSE}
WEIGHTS    <- c(0.50, 0.30, 0.20)
P          <- 600
TOL        <- 1e-5
MAX_ROUNDS <- 3000

NODES <- list(node1, node2, node3)
LAMS  <- c(lam1, lam2, lam3)

run_federated <- function(nodes, lams, weights, E,
                          tol = TOL, max_rounds = MAX_ROUNDS) {
  n_nodes      <- length(nodes)
  beta         <- numeric(P)
  loss_history <- numeric(0)

  for (t in seq_len(max_rounds)) {
    beta_old <- beta

    local_betas <- lapply(seq_len(n_nodes), function(j) {
      node_update(nodes[[j]], beta, lams[j], E)
    })

    beta <- Reduce("+", Map("*", weights, local_betas))

    global_losses <- sapply(seq_len(n_nodes), function(j) {
      node_loss_report(nodes[[j]], beta, lams[j])
    })
    loss_history <- c(loss_history, sum(weights * global_losses))

    delta <- sqrt(sum((beta - beta_old)^2))
    if (delta < tol) {
      return(list(beta = beta, loss_history = loss_history, rounds = t))
    }
  }
  warning(sprintf("E=%d: reached max_rounds=%d without convergence.", E, max_rounds))
  list(beta = beta, loss_history = loss_history, rounds = max_rounds)
}
```

## Run E = 1, 5, 10

```{r run-all-E, echo=FALSE, cache=TRUE}
results <- list()
for (E in c(1, 5, 10)) {
  results[[as.character(E)]] <- run_federated(NODES, LAMS, WEIGHTS, E)
}
```

---

# Step 3 — Results

## Question 1 — Optimal $(\lambda_1, \lambda_2, \lambda_3)$

Already reported in the lambda tuning table above.

$$\lambda_1 = `r round(lam1, 4)`, \qquad
  \lambda_2 = `r round(lam2, 4)`, \qquad
  \lambda_3 = `r round(lam3, 4)`$$

---

## Question 2 — Global Training Loss Trajectory

```{r plot-loss, echo=FALSE, fig.width=11, fig.height=4.5, out.width="100%"}
library(ggplot2)

loss_df <- do.call(rbind, lapply(c("1","5","10"), function(e) {
  lh <- results[[e]]$loss_history
  data.frame(
    E     = factor(paste0("E = ", e), levels = c("E = 1","E = 5","E = 10")),
    Round = seq_along(lh),
    Loss  = lh
  )
}))

col_pal <- c("E = 1" = "#2563EB", "E = 5" = "#16A34A", "E = 10" = "#DC2626")

p_lin <- ggplot(loss_df, aes(x = Round, y = Loss, colour = E, linetype = E)) +
  geom_line(linewidth = 1.1) +
  scale_colour_manual(values = col_pal) +
  scale_linetype_manual(values = c("E = 1" = "solid",
                                   "E = 5" = "dashed",
                                   "E = 10" = "dotdash")) +
  labs(title    = "Global Training Loss (Linear Scale)",
       x        = "Communication Round",
       y        = "Weighted Training Loss",
       colour   = "Local Epochs",
       linetype = "Local Epochs") +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold"))

p_log <- ggplot(loss_df, aes(x = Round, y = Loss, colour = E, linetype = E)) +
  geom_line(linewidth = 1.1) +
  scale_y_log10() +
  scale_colour_manual(values = col_pal) +
  scale_linetype_manual(values = c("E = 1" = "solid",
                                   "E = 5" = "dashed",
                                   "E = 10" = "dotdash")) +
  labs(title    = "Global Training Loss (Log Scale)",
       x        = "Communication Round",
       y        = "Weighted Training Loss (log10)",
       colour   = "Local Epochs",
       linetype = "Local Epochs") +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold"))

if (!requireNamespace("gridExtra", quietly = TRUE)) install.packages("gridExtra")
library(gridExtra)
grid.arrange(p_lin, p_log, ncol = 2)
```

**Figure 1.** Global weighted training loss (evaluated at the aggregated
$\boldsymbol{\beta}_{t+1}$) across communication rounds for
$E = 1$ (blue/solid), $E = 5$ (green/dashed), $E = 10$ (red/dot-dash).

---

## Question 3 — Number of Active Coefficients (E = 1)

```{r active-coefs, echo=FALSE}
EPSILON <- 1e-4

beta_E1   <- results[["1"]]$beta
active_E1 <- which(abs(beta_E1) > EPSILON)

active_df <- data.frame(
  Setting      = "Federated  E = 1",
  Threshold    = format(EPSILON, scientific = TRUE),
  Active       = length(active_E1),
  Total        = P,
  Sparsity_pct = sprintf("%.1f%%", 100 * (1 - length(active_E1) / P))
)

kable(active_df,
      col.names = c("Model", "Threshold $\\varepsilon$",
                    "Active coefs.", "Total $p$", "Sparsity"),
      align     = c("l","c","c","c","c"),
      caption   = "Number of active (non-zero) regression coefficients for E = 1.",
      booktabs  = TRUE) |>
  kable_styling(latex_options = c("striped","hold_position"),
                full_width = FALSE)
```

The federated model with $E = 1$ recovers
**`r length(active_E1)` active predictors** out of $p = 600$
(threshold $\varepsilon = 10^{-4}$).

---

## Question 4 — Confusion Matrix vs. Centralized Pooled Lasso

### Centralized Pooled Lasso Baseline

```{r pooled-lasso, echo=FALSE, cache=TRUE}
library(glmnet)

X_pool_tr  <- rbind(node1$X_tr,  node2$X_tr,  node3$X_tr)
y_pool_tr  <- c(node1$y_tr,  node2$y_tr,  node3$y_tr)
X_pool_val <- rbind(node1$X_val, node2$X_val, node3$X_val)
y_pool_val <- c(node1$y_val, node2$y_val, node3$y_val)

best_lam_pool <- NA_real_
best_vl_pool  <- Inf

for (lam in LAM_GRID) {
  fit  <- glmnet(X_pool_tr, y_pool_tr, alpha = 1, lambda = lam,
                 intercept = FALSE, standardize = FALSE)
  pred <- drop(predict(fit, newx = X_pool_val, s = lam))
  vl   <- 0.5 * mean((y_pool_val - pred)^2)
  if (vl < best_vl_pool) { best_vl_pool <- vl; best_lam_pool <- lam }
}

fit_pool    <- glmnet(X_pool_tr, y_pool_tr, alpha = 1,
                      lambda = best_lam_pool,
                      intercept = FALSE, standardize = FALSE)
beta_pool   <- as.numeric(coef(fit_pool))[-1]
active_pool <- which(abs(beta_pool) > EPSILON)

pool_df <- data.frame(
  Model      = "Centralized Pooled Lasso",
  Lambda     = round(best_lam_pool, 6),
  Val_Loss   = round(best_vl_pool, 4),
  Active     = length(active_pool),
  Total      = P
)

kable(pool_df,
      col.names = c("Model", "Optimal $\\lambda$", "Val. Loss (0.5$\\times$MSE)",
                    "Active coefs.", "Total $p$"),
      align     = c("l","c","c","c","c"),
      caption   = "Centralized pooled Lasso baseline summary.",
      booktabs  = TRUE) |>
  kable_styling(latex_options = c("striped","hold_position"),
                full_width = FALSE)
```

### Confusion Matrix Utility

```{r confusion-util, echo=FALSE}
confusion_vs_ref <- function(active_fed, active_ref, p = P) {
  af <- as.integer(active_fed)
  ar <- as.integer(active_ref)
  TP <- length(intersect(af, ar))
  FP <- length(setdiff(af, ar))
  FN <- length(setdiff(ar, af))
  TN <- p - length(union(af, ar))
  c(TP = TP, FP = FP, FN = FN, TN = TN)
}
```

### Confusion Matrix for E = 1

```{r cm-E1, echo=FALSE}
cm_E1 <- confusion_vs_ref(active_E1, active_pool)

prec_E1 <- cm_E1["TP"] / (cm_E1["TP"] + cm_E1["FP"])
rec_E1  <- cm_E1["TP"] / (cm_E1["TP"] + cm_E1["FN"])

cm_mat <- matrix(
  c(cm_E1["TN"], cm_E1["FP"], cm_E1["FN"], cm_E1["TP"]),
  nrow = 2, byrow = TRUE,
  dimnames = list(
    `Ground Truth (Pooled)` = c("Inactive", "Active"),
    `Federated Prediction`  = c("Inactive", "Active")
  )
)

kable(cm_mat,
      caption  = "Confusion matrix for E = 1 vs. centralized pooled Lasso (ground truth).",
      booktabs = TRUE) |>
  kable_styling(latex_options = c("striped","hold_position"),
                full_width = FALSE) |>
  add_header_above(c(" " = 1, "Federated Prediction" = 2))

metrics_E1 <- data.frame(
  E         = 1L,
  Precision = sprintf("%.1f%%", 100 * prec_E1),
  Recall    = sprintf("%.1f%%", 100 * rec_E1),
  TP        = cm_E1["TP"],
  FP        = cm_E1["FP"],
  FN        = cm_E1["FN"],
  TN        = cm_E1["TN"]
)

kable(metrics_E1,
      col.names = c("E", "Precision", "Recall", "TP", "FP", "FN", "TN"),
      align     = c("c","c","c","c","c","c","c"),
      caption   = "Precision and recall for the federated model (E = 1).",
      booktabs  = TRUE,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped","hold_position"),
                full_width = FALSE)
```

---

# Sensitivity Analysis: E = 1, 5, 10

## Summary Table

```{r sens-table, echo=FALSE}
sens_rows <- lapply(c("1","5","10"), function(e) {
  r   <- results[[e]]
  act <- which(abs(r$beta) > EPSILON)
  cm  <- confusion_vs_ref(act, active_pool)
  prec <- 100 * cm["TP"] / (cm["TP"] + cm["FP"])
  rec  <- 100 * cm["TP"] / (cm["TP"] + cm["FN"])
  data.frame(
    E          = as.integer(e),
    Rounds     = r$rounds,
    Final_Loss = round(tail(r$loss_history, 1), 4),
    n_Active   = length(act),
    TN         = cm["TN"],
    FP         = cm["FP"],
    FN         = cm["FN"],
    TP         = cm["TP"],
    Precision  = sprintf("%.1f%%", prec),
    Recall     = sprintf("%.1f%%", rec),
    row.names  = NULL
  )
})

sens_df <- do.call(rbind, sens_rows)
rownames(sens_df) <- NULL

kable(sens_df,
      col.names = c("$E$", "Rounds", "Final Loss", "\\# Active",
                    "TN", "FP", "FN", "TP", "Precision", "Recall"),
      align     = c("c","c","c","c","c","c","c","c","c","c"),
      caption   = "Sensitivity analysis: convergence and support recovery for $E = 1, 5, 10$.",
      booktabs  = TRUE) |>
  kable_styling(latex_options = c("striped","hold_position"),
                full_width = FALSE)
```

## Confusion Matrix Heatmaps

```{r cm-heatmaps, echo=FALSE, fig.width=7, fig.height=4, out.width="80%"}
cm_long <- do.call(rbind, lapply(c("1","5","10"), function(e) {
  r   <- results[[e]]
  act <- which(abs(r$beta) > EPSILON)
  cm  <- confusion_vs_ref(act, active_pool)
  data.frame(
    Actual    = factor(c("Inactive","Inactive","Active","Active"),
                       levels = c("Active","Inactive")),
    Predicted = factor(c("Inactive","Active","Inactive","Active"),
                       levels = c("Inactive","Active")),
    Count     = c(cm["TN"], cm["FP"], cm["FN"], cm["TP"]),
    Label     = c("TN","FP","FN","TP"),
    Panel     = paste0("E = ", e, "  (", r$rounds, " rounds)")
  )
}))

cm_long$Panel <- factor(cm_long$Panel, levels = unique(cm_long$Panel))

ggplot(cm_long, aes(x = Predicted, y = Actual, fill = Count)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(aes(label = paste0(Label, "\n", Count)),
            size = 4.5, fontface = "bold", colour = "black") +
  scale_fill_gradient(low = "#DBEAFE", high = "#1D4ED8") +
  facet_wrap(~Panel, ncol = 3) +
  labs(title = "Confusion Matrices: Federated vs. Centralized Pooled Lasso",
       x = "Federated Prediction",
       y = "Ground Truth (Pooled Lasso)") +
  theme_bw(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "right")
```

**Figure 2.** Confusion matrices for $E = 1, 5, 10$ against the centralized
pooled Lasso reference.

## Overlaid Loss Curves

```{r overlay-loss, echo=FALSE, fig.width=6, fig.height=4, out.width="70%"}
conv_rounds <- sapply(c("1","5","10"), function(e) results[[e]]$rounds)
conv_loss   <- sapply(c("1","5","10"), function(e)
                      tail(results[[e]]$loss_history, 1))
annot_df <- data.frame(
  Round = conv_rounds,
  Loss  = conv_loss,
  Label = paste0("t = ", conv_rounds),
  E     = factor(c("E = 1","E = 5","E = 10"),
                 levels = c("E = 1","E = 5","E = 10"))
)

ggplot(loss_df, aes(x = Round, y = Loss, colour = E, linetype = E)) +
  geom_line(linewidth = 1.2) +
  geom_point(data = annot_df, aes(x = Round, y = Loss),
             size = 3, shape = 21, fill = "white", stroke = 1.5) +
  geom_text(data = annot_df, aes(x = Round, y = Loss, label = Label),
            hjust = -0.18, vjust = -1.2, size = 3.6, show.legend = FALSE) +
  scale_colour_manual(values = col_pal) +
  scale_linetype_manual(values = c("E = 1" = "solid",
                                   "E = 5" = "dashed",
                                   "E = 10" = "dotdash")) +
  labs(title    = "Overlaid Global Training Loss Trajectories (E = 1, 5, 10)",
       subtitle = "Loss evaluated at aggregated beta after each communication round",
       x        = "Communication Round",
       y        = "Weighted Global Training Loss",
       colour   = "Local Epochs",
       linetype = "Local Epochs") +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom",
        plot.title    = element_text(face = "bold"),
        plot.subtitle = element_text(size = 10, colour = "grey40"))
```

**Figure 3.** Overlaid loss trajectories. Open circles mark the convergence
round for each schedule.

## Coefficient Profile Comparison

```{r coef-profiles, echo=FALSE, fig.width=6, fig.height=4, out.width="70%"}
prof_df <- do.call(rbind, lapply(c("1","5","10","pool"), function(tag) {
  if (tag == "pool") {
    b   <- beta_pool
    lbl <- sprintf("Centralized Pooled Lasso  (%d active)", length(active_pool))
  } else {
    b   <- results[[tag]]$beta
    lbl <- sprintf("Federated  E = %s  (%d active)",
                   tag, sum(abs(b) > EPSILON))
  }
  nz <- which(abs(b) > EPSILON)
  if (length(nz) == 0) return(NULL)
  data.frame(Index = nz, Coef = b[nz], Model = lbl)
}))

prof_df$Model <- factor(prof_df$Model, levels = unique(prof_df$Model))

ggplot(prof_df, aes(x = Index, y = Coef)) +
  geom_segment(aes(xend = Index, yend = 0), linewidth = 0.45,
               colour = "steelblue", alpha = 0.7) +
  geom_point(size = 1.0, colour = "steelblue4") +
  facet_wrap(~Model, ncol = 1, scales = "free_y") +
  labs(title = "Non-Zero Coefficient Profiles",
       x     = "Predictor Index (1 to 600)",
       y     = "Coefficient Value") +
  theme_bw(base_size = 11) +
  theme(plot.title  = element_text(face = "bold"),
        strip.text  = element_text(face = "bold", size = 10))
```

**Figure 4.** Stem plots of non-zero coefficients for each model.
The bottom panel (pooled Lasso) is the reference support.

---

# Discussion

## Convergence Rate

```{r convergence-table, echo=FALSE}
rounds_vec  <- sapply(c("1","5","10"), function(e) results[[e]]$rounds)
savings_pct <- round((1 - rounds_vec / rounds_vec[1]) * 100)

conv_df <- data.frame(
  E         = c(1L, 5L, 10L),
  Rounds    = as.integer(rounds_vec),
  Reduction = c("baseline",
                paste0("\u2212", savings_pct[2], "%"),
                paste0("\u2212", savings_pct[3], "%"))
)

kable(conv_df,
      col.names = c("$E$", "Rounds to convergence", "Reduction vs.\\ $E = 1$"),
      align     = c("c","c","l"),
      caption   = "Convergence speed by local epoch count.",
      booktabs  = TRUE,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped","hold_position"),
                full_width = FALSE)
```

Increasing $E$ from 1 to 10 reduces communication rounds from
**`r results[["1"]]$rounds`** to **`r results[["10"]]$rounds`**
(`r savings_pct[3]`% reduction).  Each round transmits a $p = 600$-dimensional
vector per node; fewer rounds directly reduce network bandwidth cost.

## Support Recovery: Precision / Recall Trade-off

As $E$ increases:

* **Recall decreases**: more local epochs push each node toward its own local
  Lasso minimum, causing *node drift*.  The consensus model loses sight of
  features that the pooled model recovers by leveraging all data jointly.
* **Precision changes modestly**: the active set shrinks with larger $E$ but
  still contains many false positives relative to the pooled baseline.
* **Final global loss is non-monotone in $E$**: increasing $E$ greatly reduces
  the number of communication rounds, but the final global loss does not decrease
  uniformly.  In our results, $E = 5$ yields the lowest final loss, while
  $E = 10$ converges in the fewest rounds but produces a slightly higher final
  loss than $E = 5$.  This reflects the tension between faster convergence and
  node drift: with $E = 10$, each node drifts further toward its local optimum
  before the next aggregation, which can overshoot the global consensus.

## Trade-off Summary

| Concern | Small $E$ | Large $E$ |
|---------|:---------:|:---------:|
| Communication rounds | More | Fewer |
| Support recall | Higher | Lower |
| Per-round computation | Lighter | Heavier |
| Final training loss | Slightly higher | Non-monotone ($E=5$ lowest) |
| Node drift risk | Minimal | Pronounced |

**Recommendation.**  $E = 5$ provides a good balance — substantially fewer
communication rounds than $E = 1$ while still recovering a large fraction of
the pooled model's support.  When accurate variable selection is the primary
goal, $E = 1$ most faithfully approximates the centralized solution.

---

# Session Information

```{r session-info, echo=FALSE}
si <- sessionInfo()
si_df <- data.frame(
  Item  = c("R version", "Platform", "Running under"),
  Value = c(si$R.version$version.string, si$platform, si$running)
)

kable(si_df,
      col.names = c("Item", "Value"),
      align     = c("l","l"),
      caption   = "R session information.",
      booktabs  = TRUE,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped","hold_position"),
                full_width = FALSE)
```
